---
title: "Setup Database"
author: "Ben Best"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup Database

Using PostgreSQL with PostGIS extension in [Docker](https://www.docker.com/get-started):

- [kartoza/postgis - Docker Hub](https://hub.docker.com/r/kartoza/postgis/)

- [Set Up a PostGIS Database With Docker · Alex Urquhart](https://alexurquhart.com/post/set-up-postgis-with-docker/)

### create volume `pg_data`

```
docker volume create pg_data
```

### create container `postgis`, database `am` user `admin`


On command line:

```
# set password as environment variable read in from file
DB_PASS=`cat ~/private/pg_pass`

# run docker with image
docker run --name=postgis \
  -d -e POSTGRES_USER=admin \
  -e POSTGRES_PASS=$DB_PASS \
  -e POSTGRES_DBNAME=am \
  -e ALLOW_IP_RANGE=0.0.0.0/0 \
  -p 5432:5432 \
  -v pg_data:/var/lib/postgresql \
  -v /Users/bbest/docker_pg_data:/data \
  --restart=always \
  kartoza/postgis:11.0-2.5

# check logs
docker logs postgis
```

## Connect to Database 

```{r load}
library(tidyverse)
library(here)
library(glue)
library(raster)
library(DBI)
library(RPostgreSQL)

library(devtools); load_all() # library(seaoflife)

con <- get_db_con(password=readLines("~/private/pg_pass"))
```
## Load CSVs

```{r csv_to_db, eval=F}
csv_to_db(con)
```

```{r}
spp       <- tbl(con, "spp")
cells     <- tbl(con, "cells")
spp_cells <- tbl(con, "spp_cells")
obs_cells <- tbl(con, "obs_cells")
```


## Taxonomic Groups

Reference:

- Tittensor DP, C Mora, W Jetz, HK Lotze, D Ricard, EV Berghe, B Worm (2010) [Global patterns and predictors of marine biodiversity across taxa](https://doi.org/10.1038/nature09329). _Nature_

- Primarily coastal species (Tittensor et al 2010)
    - **Coastal fishes**: `Class == "Actinopterygii" & !Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae")`
    - Non-oceanic sharks: TODO with `rredlist::rl_habitats()`
    - **Non-squid cephalopods**: `Class == "Cephalopoda" & !Order %in% c("Myopsida","Oegopsida")`
    - **Pinnipeds**: `Family %in% c("Odobenidae", "Otariidae", "Phocidae")`
    - **Corals**: `Class == "Anthozoa"` (soft/hard + anemones + gorgonians)
    - **Seagrasses**: `Order == "Alismatales"`
    - **Mangroves**: `Genus == "Avicennia" & Species == "germinans"` (n=1)
- Primarily oceanic species (Tittensor et al 2010)
    - **Tunas and billfishes**: `Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae")`
    - Oceanic sharks: TODO with `rredlist::rl_habitats()`
    - **Squids**:  `Order %in% c("Myopsida","Oegopsida")`
    - **Cetaceans**: `Order == "Artiodactyla"`
    - **Euphausiids**: `Order == "Euphausiacea"`
    - Foraminifera: n=0 for Phylum=Retaria
- Others (Phylum - Class):
    - Arthropoda
        - **Crustaceans**: `Class %in% c("Malacostraca","Ostracoda","Branchiopoda","Cephalocarida","Maxillopoda")`
        - **Sea spiders**: `Class == "Pycnogonida"`
        - skipping: Arachnida (1), Merostomata (1)
    - Mollusca
        - **Gastropods**: `Class == "Gastropoda"`
        - **Bivalves**:  `Class == "Bivalvia"`
        - **Chitons**: `Class == "Polyplacophora"`
        - Cephalopods
        - Squids
    - Chordata
        - **Tunicates**: `Class %in% c("Ascidiacea","Thaliacea")`
        - **Sharks**: `Class == "Elasmobranchii"` (actually Class=Chondrichthyes, Subclass=Elasmobranchii)
        - **Reptiles**: `Class == "Reptilia"` crocodiles, sea snakes & sea turtles
    - **Echinoderms**: `Phylum == "Echinodermata"`
    - **Sponges**: `Phylum == "Porifera"`
    - **Worms**: `Phylum == "Annelida"`
    - Cnidaria
        - **Hydrozoans**: `Class == "Hydrozoa"`

```{r AquaMaps groups, eval=F}
if (!file.exists(spp_grouped_csv)){
  spp <- suppressMessages({ read_csv(spp_csv) })
  
  # spp %>% filter(is.na(group)) %>% pull(Class) %>% table() %>% sort()
  # spp %>% filter(Class == "Reptilia") %>% View() %>% pull(Class) %>% table() %>% sort
  #  
  # # https://en.wikipedia.org/wiki/Mangrove
  # spp %>% filter(Family %in% c("Acanthaceae","Arecaceae","Bombacaceae","Cyperaceae","Euphorbiaceae","Lecythidaceae","Lythraceae","Meliaceae","Myrtaceae","Pellicieraceae","Plumbaginaceae","Primulaceae","Pteridaceae","Rubiaceae","Sterculiaceae"))

  spp <- spp %>% 
    mutate(
      group = case_when(
        Class == "Actinopterygii" & !Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae") ~ "Coastal fishes",
        Class == "Cephalopoda" & !Order %in% c("Myopsida","Oegopsida") ~ "Non-squid cephalopods",
        Family %in% c("Odobenidae", "Otariidae", "Phocidae") ~ "Pinnipeds",
        Class == "Anthozoa" ~ "Corals",
        Order == "Alismatales" ~ "Seagrasses",
        Genus == "Avicennia" & Species == "germinans" ~ "Mangroves",
        Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae") ~ "Tunas & billfishes",
        Order %in% c("Myopsida","Oegopsida") ~ "Squids",
        Order == "Artiodactyla" ~ "Cetaceans",
        Order == "Euphausiacea" ~ "Euphausiids",
        Class %in% c("Malacostraca","Ostracoda","Branchiopoda","Cephalocarida","Maxillopoda") ~ "Crustaceans",
        Class == "Pycnogonida" ~ "Sea spiders",
        Class == "Gastropoda" ~ "Gastropods",
        Class == "Bivalvia" ~ "Bivalves",
        Class == "Polyplacophora" ~ "Chitons",
        Class %in% c("Ascidiacea","Thaliacea") ~ "Tunicates",
        Class == "Elasmobranchii" ~ "Sharks",
        Class == "Reptilia"~ "Reptiles",
        Phylum == "Echinodermata"~ "Echinoderms",
        Phylum == "Porifera" ~ "Sponges",
        Phylum == "Annelida" ~ "Worms",
        Class == "Hydrozoa" ~ "Hydrozoans")) %>% 
    select(-X11, -X12)
  write_csv(spp, spp_grouped_csv)
}
spp <- suppressMessages({ read_csv(spp_grouped_csv) })

# show AquaMap species counts by Phylum and Class ----
cat("\nspp$Phylum\n")
table(spp$Phylum)
cat("\nspp$Class\n")
table(spp$Class)
cat("\nspp$Class sorted by count\n")
table(spp$Class) %>% sort()
cat("\nspp$Class not in group\n")
spp %>% filter(is.na(group)) %>% pull(Class) %>% table() %>% sort()
# AquaMap species counts by Phylum and Class

# iterate over groups ----
groups <- spp %>%
  group_by(group) %>% 
  summarize(
    n = n()) %>% 
  arrange(n) %>%  #View()
  filter(!is.na(group)) %>%
  pull(group)
#groups <- sort(unique(spp$group))
#for (i in seq_along(groups)){ # i = 3
for (i in 18:length(groups)){ # i = 3
  grp       <- groups[i]
  spp_grp   <- filter(spp, group == !!grp) %>% arrange(Genus, Species)
  grp_fname <- grp %>% str_replace_all("[ -]", ".") %>% str_to_lower()
  cat(glue("group {sprintf('%02d',i)} of {length(groups)}: {grp} (n_spp = {format(nrow(spp_grp), big.mark = ',')}) - {Sys.time()}\n\n"))
  
  spp_grp_grd         <- glue("{dir_local}/spp_{grp_fname}.grd")
  spp_grp_ext_csv     <- glue("{dir_spp}/../groups/spp_{grp_fname}_ext.csv")
  spp_grp_ext_grd_csv <- glue("{dir_spp}/../groups/spp_{grp_fname}_ext_grd.csv")
  grp_be_tif  <- glue("{dir_spp}/../groups/spp_{grp_fname}_be.tif")  # binary, extinction
  #grp_rem_tif <- glue("{dir_spp}/groups/spp_{grp_fname}_rem.tif") # relative, extinction, endemism
  grp_rel_tif <- glue("{dir_spp}/../groups/spp_{grp_fname}_rel.tif") # relative, extinction, log(endemism)
  
  if (!file.exists(spp_grp_ext_csv)){
    cat(glue("  fetching spp_grp IUCN categories - {Sys.time()}\n", .trim=F))
    
    spp_grp <- spp_grp %>%
      mutate(
        genus_species = glue("{Genus} {Species}") %>% as.character(),
        #iucn_category = map_chr(genus_species, get_iucn_category),
        # parallelize with future:: 3X faster!
        iucn_future   = map(genus_species, ~future(get_iucn_category(.x))), 
        iucn_category = map_chr(iucn_future, ~value(.x)),
        iucn_weight   = iucn_weights[iucn_category],
        tif           = glue("{dir_spp}/{SPECIESID}.tif"),
        tif_ext       = glue("{dir_spp}/{SPECIESID}_ext.tif"),
        stack_name    = str_replace_all(SPECIESID, "-", ".")) %>% 
      select(-iucn_future)
    write_csv(spp_grp, spp_grp_ext_csv)
  }
  spp_grp <- suppressMessages({ read_csv(spp_grp_ext_csv) })
  # quick fix
  spp_grp <- spp_grp %>%
    mutate(
      tif_ext = glue("{dir_spp}/{SPECIESID}_ext.tif"))
  write_csv(spp_grp, spp_grp_ext_csv)
  spp_grp <- suppressMessages({ read_csv(spp_grp_ext_csv) })
  
  cat(glue("  extend species rasters to globe - {Sys.time()}\n", .trim=F))
  #sapply(spp_grp$tif, function(tif, tif_ext) raster(tif, crs=leaflet:::epsg4326) %>% extend(r_na))
  extend_tif <- function(tif, tif_ext){
    if (file.exists(tif_ext)) return(F)
    raster(tif, crs=leaflet:::epsg4326) %>% 
      extend(r_na) %>% 
      writeRaster(tif_ext)
    return(T)
  }
  # fs::file_delete(spp_grp$tif_ext)
  future_tifs <- map2(spp_grp$tif, spp_grp$tif_ext, ~future(extend_tif(.x, .y)))
  future_vals <- map_lgl(future_tifs, ~value(.x))

  # TODO: spp raster here
  
  if (any(!file.exists(spp_grp_ext_grd_csv, grp_be_tif, grp_rel_tif))){
    cat(glue("  read raster stack of species - {Sys.time()}\n", .trim=F))
    #s_g <- stack(spp_grp_grd)
    s_g <- stack(spp_grp$tif_ext)
    names(s_g) <- spp_grp$SPECIESID
  }
  
  if (!file.exists(spp_grp_ext_grd_csv)){
    cat(glue("  rescaling with cellStats - {Sys.time()}\n", .trim=F))

    # get stats on sum RES and area
    spp_grp <- spp_grp %>%
      mutate(
        sum_RES        = map_dbl(stack_name, get_spp_stat, s_g, "sum_RES"),
        sum_area_km2   = map_dbl(stack_name, get_spp_stat, s_g, "sum_area_km2"),
        rescale_RES    = rescale(1/sum_RES),
        logrescale_RES = rescale(1/log(sum_RES)))
    
    write_csv(spp_grp, spp_grp_ext_grd_csv)
  }
  spp_grp <- suppressMessages({ read_csv(spp_grp_ext_grd_csv) })
  #table(spp_grp$iucn_weight)

  # spp_grp %>%
  #   select(SPECIESID, FBname, genus_species, 
  #          iucn_category, iucn_weight, 
  #          sum_area_km2, sum_RES, rescale_RES, logrescale_RES) %>%
  #   DT::datatable()
  # table(spp_g$iucn_category, useNA = "ifany")

  if (!file.exists(grp_be_tif)){
    cat(glue("  sum_am_stack(binary, extinction) - {Sys.time()}\n", .trim=F))
    r_be  <- sum_am_stack(
      s_g, spp_grp, do_rel=F, do_ext=T, do_rng=F, do_sca=T, binary_threshold = 0.4)
    writeRaster(r_be, grp_be_tif)
  }  
  #r_be <- raster(grp_be_tif)
  if (!file.exists(grp_rel_tif)){
    cat(glue("  sum_am_stack(relative, extinction, log(endemism)) - {Sys.time()}\n", .trim=F))
    r_rel <- sum_am_stack(
      s_g, spp_grp, do_rel=T, do_ext=T, do_rng=T, log_rng=T, do_sca=T)
    writeRaster(r_rel, grp_rel_tif)
  }
  #r_rel <- raster(grp_rel_tif)

  # cat(glue("  map_am_sum() - {Sys.time()}\n", .trim=F))
  # map_am_sum(r_be,  glue("{grp}: Binary<br> * Extinction"))
  # map_am_sum(r_rel, glue("{grp}: Relative<br> * Extinction<br> * log(Endemism)"))
}
```


## Vignette Documentation...

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
